Manuscript analyses

### 2025-02-12 cilta_cel/Manuscript analyses.Rmd
Add additional analyses

@@

## Table 3 A-B. Cytopenias following ide-cel
Evaluable patients (who survived and did not progressed) are included for each time points.
```{r Table 3}
tbl1 <- cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  select("Leukopenia" = Leukopenia_apheresis, 
         "Grade3+ Leukopenia" = gr3_Leukopenia_apheresis, 
         "Neutropenia" = Neutropenia_apheresis, 
         "Grade3+ Neutropenia" = gr3_Neutropenia_apheresis, 
         "Anemia" = Anemia_apheresis, 
         "Grade3+ Anemia" = gr3_Anemia_apheresis, 
         "Thrombocytopenia" = Thrombocytopenia_apheresis, 
         "Grade3+ Thrombocytopenia" = gr3_Thrombocytopenia_apheresis,
         "Any Cytopenia" = cytopenia_apheresis,
         "Grade3+ Cytopenia" = gr3_cytopenia_apheresis
  ) %>% 
  tbl_summary(type = list(everything() ~ "dichotomous"),
              digits = all_continuous() ~ 1,
              missing = "no"
  ) %>% 
  bold_labels() %>% add_stat_label()

@@

## Table S8. Details of infections after ide-cel
Patients can be counted multiple times if multiple infections.
```{r infection details 30 first days}
tbl1 <- cilta %>% 
  filter(time_from_infusion_to_infection_days <= 30) %>% 
  select(infection_type_bacterial_viral_fungal_parasitic) %>% 
  tbl_summary() %>%
  remove_row_type(variables = infection_type_bacterial_viral_fungal_parasitic,
                  type= "level",
                  level_value = c("Fungal", "Viral")) %>%
  remove_row_type(infection_type_bacterial_viral_fungal_parasitic, type = "header") %>%
  add_stat_label(label = infection_type_bacterial_viral_fungal_parasitic ~ c("n (%)"), location = "column") %>%
  modify_column_indent(columns = label, undo = TRUE) %>% 
  bold_levels() %>% 
  modify_header(update = all_stat_cols() ~ "**{level}**, N = {n}") %>%
  modify_footnote(update = everything() ~ NA) %>%
  modify_spanning_header(all_stat_cols() ~ "**Infections in the first 30 days**")

@@

```{r Figure 2. Inflammatory markers after ide-cel, fig.height=12}
LDH <- fig_dat %>% 
  select(pt_identifier, starts_with("ldh_")) %>% 
  pivot_longer(cols = -c(pt_identifier)) %>% 
  mutate(name = str_remove(name, "_k_ul|_k_u_l"),
         day = case_when(
           str_detect(name, "apheresis")            ~ "-28",
           str_detect(name, "pretreatment")         ~ "-28",
           str_detect(name, "day")                  ~ ((str_match(name, "([0-9]+)"))[,2])
         ), 
         day = case_when(
           day == 5                                 ~ "-5",
           TRUE                                     ~ day
         ),
         day = factor(day, levels = c("-28", "-5", "0", 
                                      "7", "14", "21", "30", "90", "180"))) %>% 
  ggplot(aes(x=day, y=value)) + 
  geom_boxplot() + 
  xlab("Days") + ylab("LDH represented with a log2 scale") + 
  theme_classic() + ggtitle("LDH") + 
  scale_y_continuous(trans = "log2") +
  stat_compare_means(label.x = 1.2 
                     # aes(label = paste0("p = ", ..p.format..))
  #                  aes(
  # label=ifelse(..p.format.. == 1e-04, "p<0.001", "Other")
  )

@@

## Figure S3. Viral titers after cilta-cel
Focus on MCC patients / I didn't use day 180 as they are less values but I can add it if you want.
```{r Figure S1. Viral titers after ide-cel}
fig_dat <- cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE)
moffitt_subset <- fig_dat %>% filter(site == "Moffitt")
```

@@

```{r Figure S2. Pneumococcal titers after ide-cel, fig.width=15, fig.height=12}
library(ComplexHeatmap)
mat <- moffitt_subset %>% 
  select(pt_identifier, 
         intersect(starts_with("pneumo"), ends_with("change"))) %>% 
  rename(PolyAb3 = pneumo_poly_ab3_change, PolyAb6 = pneumo_poly_ab6_change, 
         PolyAb14 = pneumo_poly_ab14_change, PolyAb19 = pneumo_poly_ab19_change,
         PolyAb23 = pneumo_poly_ab23_change, Serotype1 = pneumo_serotype1_ig_g_change,
         Serotype4 = pneumo_serotype_4_ig_g_change, Serotype5 = pneumo_serotype5_ig_g_change,
         Serotype7F = pneumo_serotype7f_ig_g_change, Serotype8 = pneumo_serotype8_ig_g_change,
         Serotype9N = pneumo_serotype9n_ig_g_change, Serotype9V = pneumo_serotype9v_ig_g_change,
         Serotype12F = pneumo_serotype12f_ig_g_change, Serotype18C = pneumo_serotype18c_ig_g_change
         ) %>% 
  drop_na() %>% 
  mutate(new_sort =
           c(4,3,12,31,13,14,9,15,5,6,8,10,32,16,18,17,19,30,2,1,20,21,40,22),
         .before = 2) %>%

>>>>

```{r death}
cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  mutate(os_event = case_when(
    os_event == 1          ~ "Died",
    os_event == 0          ~ "Alive"
  )) %>% 
  select(os_event) %>% 
  tbl_summary() %>% 
  bold_labels() %>% add_stat_label()

cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  filter(os_event == 1) %>% 
  select(`cause of death` = cause_of_death_myeloma_toxicitiy_unrelated) %>% 
  tbl_summary() %>% 
  bold_labels() %>% add_stat_label()
```

@@

## Table 3 A-B. Cytopenias following cilta-cel
Evaluable patients (who survived and did not progressed) are included for each time points.
```{r Table 3}
tbl1 <- cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  select("Leukopenia" = Leukopenia_apheresis, 
         "Grade3+ Leukopenia" = gr3_Leukopenia_apheresis, 
         "Neutropenia" = Neutropenia_apheresis, 
         "Grade3+ Neutropenia" = gr3_Neutropenia_apheresis, 
         "Anemia" = Anemia_apheresis, 
         "Grade3+ Anemia" = gr3_Anemia_apheresis, 
         "Thrombocytopenia" = Thrombocytopenia_apheresis, 
         "Grade3+ Thrombocytopenia" = gr3_Thrombocytopenia_apheresis,
         "Any Cytopenia" = cytopenia_apheresis,
         "Grade3+ Cytopenia" = gr3_cytopenia_apheresis
  ) %>% 
  tbl_summary(type = list(everything() ~ "dichotomous"),
              digits = all_continuous() ~ 1,
              missing = "no"
  ) %>% 
  bold_labels() %>% add_stat_label()

@@

## Table S2. severe infection
```{r}
cilta %>% 
  filter(time_from_infusion_to_infection_days <= 30) %>% 
  select(any_sevinfection) %>% 
  tbl_summary() %>% 
  bold_labels() %>% add_stat_label() %>%
  modify_spanning_header(all_stat_cols() ~ "**Severe infections in the first 30 days**")

cilta %>% 
  filter(time_from_infusion_to_infection_days > 30 &
           time_from_infusion_to_infection_days <= 100) %>% 
  select(any_sevinfection) %>% 
  tbl_summary() %>% 
  bold_labels() %>% add_stat_label() %>%
  modify_spanning_header(all_stat_cols() ~ "**Severe infections day 31 – 100**")

cilta %>% 
  filter(time_from_infusion_to_infection_days > 100) %>% 
  select(any_sevinfection) %>% 
  tbl_summary() %>% 
  bold_labels() %>% add_stat_label() %>%
  modify_spanning_header(all_stat_cols() ~ "**Severe infections after day 100**")

## Table S8. Details of infections after cilta-cel
Patients can be counted multiple times if multiple infections.
```{r infection details 30 first days}
tbl1 <- cilta %>% 
  filter(time_from_infusion_to_infection_days <= 30) %>% 
  select(infection_type_bacterial_viral_fungal_parasitic) %>% 
  tbl_summary() %>%
  remove_row_type(variables = infection_type_bacterial_viral_fungal_parasitic,
                  type= "level",
                  level_value = c("Fungal", "Viral")) %>%
  remove_row_type(infection_type_bacterial_viral_fungal_parasitic, type = "header") %>%
  add_stat_label(label = infection_type_bacterial_viral_fungal_parasitic ~ c("n (%)"), location = "column") %>%
  modify_column_indent(columns = label, undo = TRUE) %>% 
  bold_levels() %>% 
  modify_header(update = all_stat_cols() ~ "**{level}**, N = {n}") %>%
  modify_footnote(update = everything() ~ NA) %>%
  modify_spanning_header(all_stat_cols() ~ "**Infections in the first 30 days**")

@@

```{r Figure 2. Inflammatory markers after cilta-cel, fig.height=12}
LDH <- fig_dat %>% 
  select(pt_identifier, starts_with("ldh_")) %>% 
  pivot_longer(cols = -c(pt_identifier)) %>% 
  mutate(name = str_remove(name, "_k_ul|_k_u_l"),
         day = case_when(
           str_detect(name, "apheresis")            ~ "-28",
           str_detect(name, "pretreatment")         ~ "-28",
           str_detect(name, "day")                  ~ ((str_match(name, "([0-9]+)"))[,2])
         ), 
         day = case_when(
           day == 5                                 ~ "-5",
           TRUE                                     ~ day
         ),
         day = factor(day, levels = c("-28", "-5", "0", 
                                      "7", "14", "21", "30", "90", "180"))) %>% 
  ggplot(aes(x=day, y=value)) + 
  geom_boxplot() + 
  xlab("Days") + ylab("LDH represented with a log2 scale") + 
  theme_classic() + ggtitle("LDH") + 
  scale_y_continuous(trans = "log2") +
  stat_compare_means(label.x = 1.2 
                     # aes(label = paste0("p = ", ..p.format..))
  #                  aes(
  # label=ifelse(..p.format.. == 1e-04, "p<0.001", "Other")
  )

@@

## Figure S3. Viral titers after cilta-cel
Focus on MCC patients / I didn't use day 180 as they are less values but I can add it if you want.
```{r Figure S1. Viral titers after cilta-cel}
fig_dat <- cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE)
moffitt_subset <- fig_dat %>% filter(site == "Moffitt")
```

@@

```{r Figure S2. Pneumococcal titers after cilta-cel, fig.width=15, fig.height=12}
library(ComplexHeatmap)
mat <- moffitt_subset %>% 
  select(pt_identifier, 
         intersect(starts_with("pneumo"), ends_with("change"))) %>% 
  rename(PolyAb3 = pneumo_poly_ab3_change, PolyAb6 = pneumo_poly_ab6_change, 
         PolyAb14 = pneumo_poly_ab14_change, PolyAb19 = pneumo_poly_ab19_change,
         PolyAb23 = pneumo_poly_ab23_change, Serotype1 = pneumo_serotype1_ig_g_change,
         Serotype4 = pneumo_serotype_4_ig_g_change, Serotype5 = pneumo_serotype5_ig_g_change,
         Serotype7F = pneumo_serotype7f_ig_g_change, Serotype8 = pneumo_serotype8_ig_g_change,
         Serotype9N = pneumo_serotype9n_ig_g_change, Serotype9V = pneumo_serotype9v_ig_g_change,
         Serotype12F = pneumo_serotype12f_ig_g_change, Serotype18C = pneumo_serotype18c_ig_g_change
         ) %>% 
  drop_na() %>% 
  mutate(new_sort =
           c(4,3,12,31,13,14,9,15,5,6,8,10,32,16,18,17,19,30,2,1,20,21,40,22),
         .before = 2) %>%
